Automatic Detection of Bad Programming Habits in Scratch: A Preliminary Study

### 2014-04-20 paperFIE_en.tex
Introduction: add pararaph with the goal of this paper. Modify some expressions.

\begin{abstract}
%This paper describes a method for automatic detection of two bad programming habits that students learning to program with Scratch incur, such as the repetition of code and object naming, and raises ideas to try to avoid such situations. FIXME: some methodology, results and conclusions should be included in the abstract.
Using Scratch as a tool to teach Computer Science or develop Computational Thinking from other disciplines is increasingly common in all levels of education, and it is possible to find well-documented success case studies from primary school to university experiences.

However, there are reports indicating that students learning to program in this environment show certain habits that are contrary to the basic programming recommendations. In our work as instructors with high school students learning to program with Scratch, we also detected some common bad practices, such as the repetition of code and object naming, and this study focuses on checking whether these issues can be found in the projects shared on the community website.

To test this hypothesys, we downloaded 100 projects from the Scratch website and analyzed them with two plug-ins we developed to automatically detect these bad practices. The plug-ins extend the functionality of the tool Hairball, a static code analyzer for Scratch projects. The obtained results indicate that, in general, projects in the repository also incur in the investigated malpractices.

Along with suggestions for future work, some ideas that might assist to avoid such situations are proposed in the conclusions of the paper.

@@

However, it is also possible to find studies that have detected some bad programming habits in students learning to program in this environment~\cite{meerbaum2011habits}. In this line, in our work as instructors we have detected bad habits despite having insisted our students to avoid them. As these habits are contrary to the basic programming recommendations, our study tries to find out if they are also common in the projects shared in the Scratch repository, and propose some ideas to try to avoid such situations.

FIXME: the goal of this paper is to... 
% ¿No estaría explicado justo en el párrafo de encima? ¿Habría que explicarlo más detalladamente?

@@

In our work as instructors with high school students, we found that, despite our constant insistence and warnings to try to avoid them, many students show two bad habits in a regular basis that have to do with object naming and code repetition.

@@

This paper shows the preliminary results of a study regarding two bad programming habits we have detected in our work as instructors with high school students learning to program with Scratch. In order to check if these bad habits, which have to do with object naming and code repetition, are also commonly found in the projects shared in the community repository, we downloaded 100 proyects and analyzed them with two plug-ins we developed for Hairball, detecting that most of the inspected proyects, 79\% and 62\%  respectively, fell into these issues.

>>>>

\begin{abstract}
%This paper describes a method for automatic detection of two bad programming habits that students learning to program with Scratch incur, such as the repetition of code and object naming, and raises ideas to try to avoid such situations. FIXME: some methodology, results and conclusions should be included in the abstract.
Using Scratch as a tool to teach Computer Science or develop Computational Thinking from other disciplines is increasingly common in all levels of education, and it is possible to find well-documented case studies from primary school to university.

However, there are reports indicating that students learning to program in this environment show certain habits that are contrary to the basic programming recommendations. In our work as instructors with high school students learning to program with Scratch, we have detected other bad practices, such as the repetition of code and object naming, that are repeated on a regular basis, and this study focuses on checking whether these issues are also present at the projects shared on the community website.

To test this hypothesys, we downloaded 100 projects from the Scratch website and analyzed them with two plug-ins we developed to automatically detect these bad practices. The plug-ins extend the functionality of the tool Hairball, a static code analyzer of Scratch projects.

The obtained results indicate that, in general, projects in the repository also incur in the investigated malpractices. Along with suggestions for future work, some ideas that might assist to avoid such situations are proposed in the conclusions of the paper.

@@

However, it is also possible to find studies that have detected some bad programming habits in students learning to program in this environment~\cite{meerbaum2011habits}. In this line, in our work as instructors we have also found some bad habits despite having insisted our students to avoid them. As these habits are contrary to the basic programming recommendations, our study tries to find out if they are also common in the projects shared in the Scratch repository.

%FIXME: the goal of this paper is to... 
The goal of this paper, therefore, is to determine whether these bad habits are common practices in the Scratch community and propose some ideas to try to avoid such situations. In order to assist evaluators to detect these malpractices we developed two plug-ins that automate their detection by extending the functionality of Hairball, a static code analyzer of Scratch projects.

@@

In our work as instructors with high school students, we found that, despite our constant insistence and warnings to try to avoid them, many students show two bad habits on a regular basis that have to do with object naming and code repetition.

@@

This paper shows the preliminary results of a study regarding two bad programming habits we have detected in our work as instructors with high school students learning to program with Scratch. In order to check if these bad habits, which have to do with object naming and code repetition, are also commonly found in the projects shared in the community repository, we downloaded 100 proyects and analyzed them with two plug-ins we developed for Hairball, detecting that most of the inspected proyects, 79\% and 62\%  respectively, fall into these issues.
